\textbf{SPECFEM3D}\\ \textbf{Cartesian}\\ \textbf{User Manual}

### 2022-11-21 doc/USER_MANUAL/06_fault_sources.tex
updates doc (newlines for better conversion with pandoc)

\item Seismograms at stations in the bulk (out of the fault plane) given
in \texttt{DATA/STATIONS}.
\item Rupture time files are named $\mathtt{Rupture\_time\_FAULT}$-$\mathtt{id}$.
One file is generated for each fault. The files are ascii and start
with a header (12 lines long) followed by a data block with the following
format, one line per fault node:

Column 1 = horizontal coordinate, distance along strike (m)

Column 2 = vertical coordinate, distance down-dip (m)

Column 3 = rupture time (s)

>>>>

\item Seismograms at stations in the bulk (out of the fault plane) given
in \texttt{DATA/STATIONS}.
\item Rupture time files are named $\mathtt{Rupture\_time\_FAULT}$-$\mathtt{id}$.
One file is generated for each fault. The files are ascii and start
with a header (12 lines long) followed by a data block with the following
format, one line per fault node:\newline

Column 1 = horizontal coordinate, distance along strike (m)\newline

Column 2 = vertical coordinate, distance down-dip (m)\newline

Column 3 = rupture time (s)\newline

### 2022-11-21 doc/USER_MANUAL/07_adjoint_simulations.tex
updates doc (newlines for better conversion with pandoc)

Instead of the kernels for the isotropic wave speeds, you can also
compute the kernels for the 21 independent components $C_{IJ},\, I,J=1,...,6$
(using Voigt's notation) of the elastic tensor in the cartesian coordinate
system. This is done by setting \texttt{ANISOTROPIC\_KL} \texttt{=}
\texttt{.true.} in \texttt{DATA/Par\_file} before compiling the package.
The definition of the parameters $C_{IJ}$ in terms of the corresponding
components $c_{ijkl},ijkl,i,j,k,l=1,2,3$ of the elastic tensor in
cartesian coordinates follows \citet{ChTr07}. The 21 anisotropic
kernels are saved in the \texttt{LOCAL\_PATH} in one file with the
name of \texttt{proc??????\_cijkl\_kernel.bin} (with \texttt{proc??????}
the processor number). The output kernels correspond to absolute perturbations
$\delta C_{IJ}$ of the elastic parameters and their unit is in $s/GPa/km^{3}$.
For consistency, the output density kernels with this option turned
on are for a perturbation $\delta\rho$ (and not $\frac{\delta\rho}{\rho}$)
and their unit is in s / (kg/m$^{3}$) / km$^{3}$. These `primary'
anisotropic kernels can then be combined to obtain the kernels for
different parameterizations of anisotropy. This can be done, for example,
when combining the kernel files from slices into one mesh file (see
Section~\ref{sec:Finite-Frequency-Kernels}).

If \texttt{ANISOTROPIC\_KL} \texttt{=} \texttt{.true.} by additionally
setting \texttt{SAVE\_TRANSVERSE\_KL} \texttt{=} \texttt{.true.} in \texttt{DATA/Par\_file}
the package will save anisotropic kernels parameterized as velocities
related to transverse isotropy based on the the Chen and Tromp parameters
\citet{ChTr07}. The kernels are saved as relative perturbations for
horizontal and vertical P and S velocities, $\alpha_{v},\alpha_{h},\beta_{v},\beta_{h}$.
Explicit relations can be found in appendix B. of \citet{SiLiTrTr07b}

The anisotropic kernels are only currently available for CPU mode.

\end{enumerate}
In general, the three steps need to be run sequentially to assure
proper access to the necessary files. If the simulations are run through
some cluster scheduling system (e.g., LSF), and the forward simulation
and the subsequent kernel simulations cannot be assigned to the same
set of computer nodes, the kernel simulation will not be able to access
the database files saved by the forward simulation. Solutions for
this dilemma are provided in Chapter~\ref{cha:Scheduler}. Visualization
of the finite-frequency kernels is discussed in Section~\ref{sec:Finite-Frequency-Kernels}.

>>>>

Instead of the kernels for the isotropic wave speeds, you can also
compute the kernels for the 21 independent components $C_{IJ},\, I,J=1,...,6$
(using Voigt's notation) of the elastic tensor in the cartesian coordinate
system. This is done by setting \texttt{ANISOTROPIC\_KL} \texttt{=}
\texttt{.true.} in \texttt{DATA/Par\_file} before compiling the package.
The definition of the parameters $C_{IJ}$ in terms of the corresponding
components $c_{ijkl},ijkl,i,j,k,l=1,2,3$ of the elastic tensor in
cartesian coordinates follows \citet{ChTr07}. The 21 anisotropic
kernels are saved in the \texttt{LOCAL\_PATH} in one file with the
name of \texttt{proc??????\_cijkl\_kernel.bin} (with \texttt{proc??????}
the processor number). The output kernels correspond to absolute perturbations
$\delta C_{IJ}$ of the elastic parameters and their unit is in $s/GPa/km^{3}$.
For consistency, the output density kernels with this option turned
on are for a perturbation $\delta\rho$ (and not $\frac{\delta\rho}{\rho}$)
and their unit is in s / (kg/m$^{3}$) / km$^{3}$. These `primary'
anisotropic kernels can then be combined to obtain the kernels for
different parameterizations of anisotropy. This can be done, for example,
when combining the kernel files from slices into one mesh file (see
Section~\ref{sec:Finite-Frequency-Kernels}).\newline

If \texttt{ANISOTROPIC\_KL} \texttt{=} \texttt{.true.} by additionally
setting \texttt{SAVE\_TRANSVERSE\_KL} \texttt{=} \texttt{.true.} in \texttt{DATA/Par\_file}
the package will save anisotropic kernels parameterized as velocities
related to transverse isotropy based on the the Chen and Tromp parameters
\citet{ChTr07}. The kernels are saved as relative perturbations for
horizontal and vertical P and S velocities, $\alpha_{v},\alpha_{h},\beta_{v},\beta_{h}$.
Explicit relations can be found in appendix B. of \citet{SiLiTrTr07b}\newline

\end{enumerate}

In general, the three steps need to be run sequentially to assure
proper access to the necessary files. If the simulations are run through
some cluster scheduling system (e.g., LSF), and the forward simulation
and the subsequent kernel simulations cannot be assigned to the same
set of computer nodes, the kernel simulation will not be able to access
the database files saved by the forward simulation. Solutions for
this dilemma are provided in Chapter~\ref{cha:Scheduler}. Visualization
of the finite-frequency kernels is discussed in Section~\ref{sec:Finite-Frequency-Kernels}.

### 2022-11-21 doc/USER_MANUAL/14_post_processing.tex
updates doc (newlines for better conversion with pandoc)

In many cases, the SEM synthetics are calculated and compared to data
seismograms recorded at seismic stations. Since the SEM synthetics
are accurate for a certain frequency range, both the original data
and the synthetics need to be processed before a comparison can be
made.

@@

\noindent
which has resampled the SAC files to a sampling rate of
1 seconds, cut them between 0 and 4000 seconds, transfered them into
displacement records and filtered them between 40 and 500 seconds,
picked the first P and S arrivals, and added suffix `\texttt{bp}'
to the file names.

@@

{\footnotesize
\begin{verbatim}
process_syn.pl -m CMTSOLUTION -h -a STATIONS -s 1.0 -l 0/4000 -f -t 40/500 -p -x bp SEM/*.BX?.semd
\end{verbatim}
}
\noindent
which will convolve the synthetics with a triangular source-time function
from the \texttt{CMTSOLUTION} file, convert the synthetics into SAC
format, add event and station information into the SAC headers, resample
the SAC files with a sampling rate of 1 seconds, cut them between
0 and 4000 seconds, filter them between 40 and 500 seconds with the
same filter used for the observed data, pick the first P and S arrivals,
and add the suffix `\texttt{bp}' to the file names.

More options are available for this script, such as adding time shift
to the origin time of the synthetics, convolving the synthetics with
a triangular source time function with a given half duration, etc.
Type \texttt{process\_syn.pl} without any argument for a detailed
usage.

@@

The forward and adjoint simulations generate synthetic seismograms
in the \texttt{OUTPUT\_FILES/} directory by default. For the forward
simulation, the files are named like \texttt{NT.STA.BX?.semd} for
two-column time series, or \texttt{NT.STA.BX?.semd.sac} for ASCII
SAC format, where NT and STA are the network code and station name,
and \texttt{BX?} stands for the component name. Please see the Appendix~\ref{cha:Coordinates}
and \ref{cha:channel-codes} for further details.

The adjont simulations generate synthetic seismograms with the name
\texttt{NT.S?????.S??.sem} (refer to Section \ref{sec:Adjoint-simulation-sources}
for details). The kernel simulations output the back-reconstructed
synthetic seismogram in the name \texttt{NT.STA.BX?.semd}, mainly
for the purpose of checking the accuracy of the reconstruction. Refer
to Section \ref{sec:Adjoint-simulation-finite} for further details.

>>>>

In many cases, the SEM synthetics are calculated and compared to data
seismograms recorded at seismic stations. Since the SEM synthetics
are accurate for a certain frequency range, both the original data
and the synthetics need to be processed before a comparison can be
made.\newline

@@

\noindent
which has resampled the SAC files to a sampling rate of
1 seconds, cut them between 0 and 4000 seconds, transfered them into
displacement records and filtered them between 40 and 500 seconds,
picked the first P and S arrivals, and added suffix `\texttt{bp}'
to the file names.\newline

@@

{\footnotesize
\begin{verbatim}
process_syn.pl -m CMTSOLUTION -h -a STATIONS -s 1.0 -l 0/4000 -f -t 40/500 -p -x bp SEM/*.BX?.semd
\end{verbatim}
}
\noindent
which will convolve the synthetics with a triangular source-time function
from the \texttt{CMTSOLUTION} file, convert the synthetics into SAC
format, add event and station information into the SAC headers, resample
the SAC files with a sampling rate of 1 seconds, cut them between
0 and 4000 seconds, filter them between 40 and 500 seconds with the
same filter used for the observed data, pick the first P and S arrivals,
and add the suffix `\texttt{bp}' to the file names.\newline

More options are available for this script, such as adding time shift
to the origin time of the synthetics, convolving the synthetics with
a triangular source time function with a given half duration, etc.
Type \texttt{process\_syn.pl} without any argument for a detailed
usage.\newline

@@

The forward and adjoint simulations generate synthetic seismograms
in the \texttt{OUTPUT\_FILES/} directory by default. For the forward
simulation, the files are named like \texttt{NT.STA.BX?.semd} for
two-column time series, or \texttt{NT.STA.BX?.semd.sac} for ASCII
SAC format, where NT and STA are the network code and station name,
and \texttt{BX?} stands for the component name. Please see the Appendix~\ref{cha:Coordinates}
and \ref{cha:channel-codes} for further details.\newline

The adjont simulations generate synthetic seismograms with the name
\texttt{NT.S?????.S??.sem} (refer to Section \ref{sec:Adjoint-simulation-sources}
for details). The kernel simulations output the back-reconstructed
synthetic seismogram in the name \texttt{NT.STA.BX?.semd}, mainly
for the purpose of checking the accuracy of the reconstruction. Refer
to Section \ref{sec:Adjoint-simulation-finite} for further details.\newline
